---
title: "Semester Overview"
format:
  html:
    toc: true
    number-sections: true
params:
  semester: null
  semester_start: null
  required_weeks: 10
  display_weeks: 14
editor_options: 
  chunk_output_type: console
---

```{r example-params, include=FALSE}
# Quick dev compile helper
if (FALSE) {
  params <- list(
    semester = "2025.2",
    semester_start = "2025-08-25",
    required_weeks = 10,
    display_weeks = 14
  )
}
```

```{r setup-libraries, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
library(arrow)
```

```{r load-data, echo=F}
library(targets)
tar_load(d)
tar_load(d_obs)

# Filter for this semester
ds <- d |> filter(semester == params$semester)
ds_obs <- d_obs |> filter(semester == params$semester)
```

## Semester: `r params$semester`

**Estimated semester start date:** `r params$semester_start`

There are `r nrow(ds)` total observations this semester.

**Number of observers:** `r n_distinct(ds$observedby_person_id)`  
**Total observations:** `r nrow(ds)`

[Download all observations from this semester (CSV)](https://github.com/AdamWilsonLabEDU/campus_phenology/releases/download/npn-data/npn_obs_network-891_semester-`r params$semester`.csv)

[Download observer summary CSV](https://github.com/AdamWilsonLabEDU/campus_phenology/releases/download/npn-data/npn_obs_network-891_semester-`r params$semester`_observer_summary.csv)

The NNID in the table and figures below is an anonymized individual observer ID. 

## Observer Overview

```{r observer-overview, echo=F}
grade <- ds_obs |>
  mutate(observedby_person_id = as.character(observedby_person_id)) |>
  select(observedby_person_id, week_credits, percent)

ds |>
  mutate(observedby_person_id = as.character(observedby_person_id)) |>
  group_by(observedby_person_id) |>
  summarize(total_obs = n(), .groups = "drop") |>
  left_join(grade, by = "observedby_person_id") |>
  mutate(
    grade = ifelse(is.na(percent), "0%", sprintf("%.0f%%", percent)),
    link = paste0("<a href='student_", params$semester, "_", observedby_person_id, ".html'>See details</a>")
  ) |>
  select(NNID = observedby_person_id, total_obs, `Weeks credited` = week_credits, Grade = grade, link) |>
  datatable(escape = FALSE)
```


## Weekly Observations Heatmap


The figure below visualizes the number of observations made by each observer during each week of the semester. 

```{r weekly-heatmap, echo=F, fig.height=12 }
if (nrow(ds) > 0) {
  week_breaks <- seq_len(params$display_weeks)

  ds |>
    filter(semester_week >= 1, semester_week <= params$display_weeks) |>
    group_by(observedby_person_id, semester_week) |>
    summarize(total_obs = n(), .groups = "drop") |>
    ggplot(aes(x = semester_week, y = as.factor(observedby_person_id), fill = total_obs)) +
    geom_tile() +
    scale_x_continuous(
      breaks = week_breaks,
      limits = c(0.5, params$display_weeks + 0.5),
      expand = c(0, 0)
    ) +
    scale_fill_viridis_c(trans = "log1p") +
    xlab("Week of Semester") +
    ylab("Observer") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
}
```



## Phenology Observations by Species

```{r species-observations, echo=F}
if (nrow(ds) > 0) {
  ds |>
    group_by(common_name) |>
    summarize(n_obs = n()) |>
    ggplot(aes(x = reorder(common_name, n_obs), y = n_obs)) +
    geom_col() +
    coord_flip() +
    ylab("Number of Observations") +
    xlab("Species")
}
```

## How observation grades are calculated

```{r grading-dates, echo=FALSE}
semester_start <- as.Date(params$semester_start)
semester_end <- semester_start + params$required_weeks * 7 - 1
```

The estimated start of this semester is `r if (is.na(semester_start)) "not set" else format(semester_start, "%B %d, %Y")`. The grading period is the `r params$required_weeks` weeks that follow, through `r if (is.na(semester_end)) "the end of week 10" else format(semester_end, "%B %d, %Y")`.

Fall starts on the first Monday on or after August 23. Spring starts on the Monday of the week that contains January 25. Each grading week is a standard Monday–Sunday ISO week beginning on that Monday.

A student's grade uses only **Leaves** observations inside that window:

- Any Leaves visit in a Monday–Sunday week fills that week. Extra visits in the same week do not add points.
- The grade is 100 times the number of filled weeks, divided by `r params$required_weeks`. Ten filled weeks is 100%.
- Visits after week `r params$required_weeks` do not add points.

The weekly heatmap shows every observation across the full `r params$display_weeks`-week semester calendar. The grade only counts Leaves visits in the first `r params$required_weeks` weeks.


