---
title: "Student Observations"
format:
  html:
    toc: true
    number-sections: true
params:
  semester: null
  semester_start: null
  nnid: null
  required_weeks: 10
  display_weeks: 14
---

```{r example-params, include=FALSE}
# Quick dev compile helper
if (FALSE) {
  params <- list(
    semester = "2025.2",
    semester_start = "2025-08-25",
    nnid = 60171,
    required_weeks = 10,
    display_weeks = 14
  )
}
```

```{r setup-libraries, echo=F, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(arrow)
```

```{r load-data, echo=F}
library(targets)

tar_load(d)
tar_load(d_obs)

nnid_chr <- as.character(params$nnid)

# Filter for this student and semester
ds <- d |>
  filter(
    semester == params$semester,
    as.character(observedby_person_id) == nnid_chr
  )

ds_obs <- d_obs |>
  filter(
    semester == params$semester,
    as.character(observedby_person_id) == nnid_chr
  )

week_credits <- if (nrow(ds_obs) == 0) 0L else ds_obs$week_credits[[1]]
grade_percent <- if (nrow(ds_obs) == 0) 0 else ds_obs$percent[[1]]
median_date_dif <- if (nrow(ds_obs) == 0) NA_real_ else ds_obs$median_date_dif[[1]]
```

# Student Overview

## **NNID:** `r params$nnid`  

**Semester:** `r params$semester`  
**Estimated semester start date:** `r params$semester_start`  
**Total observations:** `r nrow(ds)`  
**Weeks credited:** `r week_credits` of `r params$required_weeks`

**Observation grade:** `r sprintf("%.0f%%", grade_percent)`  
**Median submission delay (days):** `r if (is.na(median_date_dif)) "NA" else sprintf("%.1f", median_date_dif)`

---

## Weekly Observations

```{r echo=F}
week_breaks <- seq_len(params$display_weeks)

ds |>
  filter(semester_week >= 1, semester_week <= params$display_weeks) |>
  group_by(semester_week) |>
  summarize(n_obs = n(), .groups = "drop") |>
  ggplot(aes(x = semester_week, y = n_obs)) +
  geom_col(fill = "steelblue") +
  scale_x_continuous(
    breaks = week_breaks,
    limits = c(0.5, params$display_weeks + 0.5),
    expand = c(0, 0)
  ) +
  xlab("Week") +
  ylab("Number of Observations")
```

---

## Observations by Plant

The number of observations per individual plant monitored by this student.

```{r observations-by-plant, echo=F}
ds |>
  group_by(plant_nickname) |>
  summarize(n_obs = n()) |>
  ggplot(aes(x = reorder(plant_nickname, n_obs), y = n_obs)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  ylab("Observations") +
  xlab("Plant Nickname")
```
