---
title: "Phenology Observations at the University at Buffalo"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: cosmo
editor: visual
execute:
  echo: false
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

```{r setup-data, echo=F, warning=FALSE, message=FALSE}
# Global options and libraries
library(tidyverse)
library(lubridate)
library(zoo)
library(sf)
library(leaflet)
library(DT)
library(arrow)
library(cowplot)

# Load  datasets (Parquet via Arrow)

data_dir <- "data/processed"

trees <- read_parquet(file.path(data_dir, "trees.parquet"))

d <- read_parquet(
file.path(data_dir, "full_data.parquet")
)


d_obs_weekly <- read_parquet(
file.path(data_dir, "weekly_observer_stats.parquet")
)

d_obs <- read_parquet(
file.path(data_dir, "semester_observer_stats.parquet")
)
```

# Overview

Phenology is the study of the timing of ecological events such as leaf-out in spring, flowering in summer, and senescence in fall. These events are sensitive indicators of climate variability and change.

Throughout the semester, students in the Earth, Environment, and Climate Lab (GEO105) at the University at Buffalo monitor trees on North Campus and upload observations to Nature’s Notebook and the USA National Phenology Network. This site summarizes those observations.

These data were last updated on `r Sys.time()`.

# Trees at the University at Buffalo

The map below shows the locations of monitored trees on UB's North Campus.

```{r tree-map, echo=F}
# Prepare tree data with observation stats
tree_obs_stats <- d %>%
  filter(!is.na(individual_id)) %>%
  group_by(individual_id) %>%
  summarize(
    n_obs = n(),
    last_obs_date = max(observation_date),
    .groups = "drop"
  )

trees_with_stats <- trees %>%
  left_join(tree_obs_stats, by = "individual_id") %>%
  mutate(
    n_obs = replace_na(n_obs, 0),
    last_obs_date = replace_na(last_obs_date, as.Date(NA))
  )

# Create popups with HTML formatting
trees_with_stats <- trees_with_stats %>%
  mutate(
    popup_html = paste0(
      "<b>Tag:</b> ", tag, "<br>",
      "<b>Species:</b> ", common_name, "<br>",
      "<b>Scientific Name:</b> <i>", species, "</i><br>",
      "<b>Observations:</b> ", n_obs, "<br>",
      "<b>Last Observation:</b> ", format(last_obs_date, "%Y-%m-%d"), "<br>",
      "<a href='generated/tree_", individual_id, ".html'>View tree details</a>"
    )
  )

icons <- awesomeIcons(
  icon = "ios-close",
  iconColor = "white",
  library = "ion",
  markerColor = trees_with_stats$color
)

leaflet(trees_with_stats) %>%
  addTiles() %>%
  addAwesomeMarkers(
    ~lon, ~lat,
    icon = icons,
    popup = ~popup_html,
    label = paste(
      "Tag:", trees_with_stats$tag, "|",
      "Species:", trees_with_stats$common_name
    )
  )
```

# Current Observation Status

To date, the program has collected `r format(nrow(d), big.mark = ",")` individual phenological observations.

## Observations by Species

```{r species-table, echo=F}
d |>
count(common_name, genus, species, name = "Observations") |>
mutate(
Species = paste(common_name, "(", genus, species, ")")
) |>
arrange(desc(Observations)) |>
select(Species, Observations) |>
collect() |>
datatable()

```

# Observations Over Time

```{r observations-timeline, echo=F}

# cumulative sum of observations ggplot
p1 <- d |>
count(observation_date) |>
collect() |>
arrange(observation_date) |>
mutate(cumulative_observations = cumsum(n)) |>
ggplot(aes(x = observation_date, y = cumulative_observations)) +
geom_line(linewidth = 1) +
labs(
x = "Date",
y = "Cumulative Observations",
) +
scale_y_continuous(labels = scales::comma)

p2 <- d |>
count(observation_date) |>
  collect() |>
ggplot(aes(observation_date, n)) +
geom_col() +
labs(
x = "Date",
y = "Daily Observations")

plot_grid(p1,p2, ncol=1, align="hv",rel_heights = c(0.7,0.3))

```

```{r observers-per-semester, echo=F, eval=F}
# Observers and Participation
## Unique Observers per Semester
d |>
distinct(semester, observedby_person_id) |>
count(semester) |>
  collect() |>
ggplot(aes(semester, n, group = 1)) +
geom_line() +
geom_point() +
  annotate(
    "text",
    x = 2020.7,
    y = 80,
    label = paste("COVID-19"),
    vjust = -1,
    size=3
  ) +
labs(
x = "Semester",
y = "Unique Observers",
title = "Student Participation Over Time"
)
```

```{r echo=F, eval=F}
## Observations per Student
d |>
count(observedby_person_id, semester) |>
  collect() |>
ggplot(aes(semester, n, group=semester)) +
  geom_boxplot() +
  labs(
    x = "Semester",
    y = "Observations per Student"
    ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

# Species-Level Sampling Effort

```{r echo=F}
d |>
filter(phenophase_description == "Leaves") |>
count(semester, common_name) |>
  collect() |>
ggplot(aes(semester, n,group=semester)) +
geom_boxplot() +
facet_wrap(~common_name, nrow = 2) +
labs(
x = "Semester",
y = "Observations per Species"
) +
theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

# Seasonal Phenology Patterns

The following plot

```{r, eval=F, echo=F}
d |>
filter(
  phenophase_description == "Leaves",
  day_of_year > 180
  ) |>
collect() |>
ggplot( aes(
            x = observation_date,
            y = intensity_category_id,
            group = common_name
          )
      ) +
geom_jitter(height = 5, alpha = 0.4) +
geom_smooth(method = "loess", span = 1, se = FALSE) +
facet_grid( year(observation_date) ~ common_name) +
labs(
    x = "Day of Year (Standardized)",
    y = "Percent Green Leaves",
    title = "Fall Leaf Senescence Across Years"
    ) +
theme(strip.text.y.right = element_text(angle = 0))
```

```{r echo=F}


d %>%
  collect() |>
  filter(phenophase_description == "Leaves",
         day_of_year > 180) %>% # Fall
  ggplot(aes(x = day_of_year, y = intensity)) +
  # Raw observations (all students)
  geom_jitter(#aes(color = as.factor(observedby_person_id)), 
              col="black",size=.1,
              width = 0, height = 5, alpha = 0.1, show.legend = FALSE) +
  # Individual tree trends
  geom_smooth(aes(group = individual_id), alpha = 0.1, size=.1,color = "blue",se=F) +
#  geom_smooth(aes(group = 1), alpha = 0.3, color = "red",se=T) +
    geom_smooth(
    method = "nls",
    formula = y ~ 100 + (B - 100) / (1 + exp((xmid - x)/scal)),
    method.args = list(start = list(B = 50, xmid = 300, scal = 1)),
    se = FALSE,
    col = "red"
  )+
  facet_grid(common_name ~ year(observation_date)) +
  labs(x = "Day of Year", y = "Percent Green Leaves",
       title = "Fall Leaf Cover Trends by Species") +
  theme_minimal()+
  coord_cartesian(ylim=c(0,100),xlim=c(255,330))

```