# Data Pipeline Status

This page summarizes the current status of the campus phenology data pipeline.

## Semester Data Table

The table below lists all semesters with available observation data, including links to per-semester pages and downloadable CSVs.

```{r semester-table, echo=FALSE}
# Load the semester index table from targets
library(targets)
tar_load(semester_index)
library(DT)
datatable(
	semester_index %>%
		dplyr::transmute(
			Semester = label,
			Status = status,
			Observers = n_observers,
			Observations = n_observations,
			Page = paste0('<a href="', page_href, '">View</a>'),
			Data = paste0('<a href="', csv_href, '">Observations CSV</a>'),
			ObserverData = paste0('<a href="', observer_csv_href, '">Observer summary CSV</a>')
		),
	escape = FALSE,
	options = list(pageLength = 15, autoWidth = TRUE)
)
```

## Pipeline Visualization

The current {targets} pipeline structure is shown below. This interactive graph helps diagnose dependencies and track progress.

```{r pipeline-graph, echo=FALSE, fig.width=12, fig.height=8}
library(targets)
tar_visnetwork(targets_only = TRUE)
```


```{r pipeline-status, echo=FALSE}
library(dplyr)
progress <- read.csv('_targets/meta/progress', sep='|')
errored <- progress %>% filter(progress == 'errored')
skipped <- progress %>% filter(progress == 'skipped')
cat('Errored targets:', paste(errored$name, collapse=', '), '\n')
cat('Skipped targets:', paste(skipped$name, collapse=', '), '\n')
```

## Key Output Files

| Output | Path |
|--------|------|
| Full data Parquet | data/processed/full_data.parquet |
| Semester observer stats Parquet | data/processed/semester_observer_stats.parquet |
| Trees Parquet | data/processed/trees.parquet |
| Weekly observer stats Parquet | data/processed/weekly_observer_stats.parquet |
| Semester CSVs | data/cache/csv/npn_obs_network-891_semester-YYYY.S.csv |
| Semester observer summary CSVs | data/cache/csv/npn_obs_network-891_semester-YYYY.S_observer_summary.csv |

## Data Freshness

```{r data-freshness, echo=FALSE}
fs <- function(path) if (file.exists(path)) file.info(path)$mtime else NA
cat('full_data.parquet:', fs('data/processed/full_data.parquet'), '\n')
cat('semester_observer_stats.parquet:', fs('data/processed/semester_observer_stats.parquet'), '\n')
cat('trees.parquet:', fs('data/processed/trees.parquet'), '\n')
cat('weekly_observer_stats.parquet:', fs('data/processed/weekly_observer_stats.parquet'), '\n')
```


## Environment Info

```{r env-info, echo=FALSE}
cat('R version:', R.version.string, '\n')
cat('targets version:', as.character(packageVersion('targets')), '\n')
cat('quarto version:', tryCatch(system('quarto --version', intern=TRUE), error=function(e) 'not found'), '\n')
cat('Loaded packages:', paste(loadedNamespaces(), collapse=', '), '\n')
```

